A programmer friendly communication framework

### 2024-03-10 2024_paper/paper/msglink_framework.tex

\begin{enumerate}
    \item While the event system provided by Socket.IO is useful for many applications, sometimes responses to events are required. This is typically named a \textbf{Remote Procedure Call} and needs to be implemented manually for every event:
    \begin{itemize}
        \item define and keep track of a separate request and response event
        \item add an identifying element to request and response data to identify which response belongs to which request and prevent mixing up responses if multiple requests are sent out in short succession
        \item keep a list of open requests to handle responses once received.
    \end{itemize}
    \item While Socket.IO provides a very clean and feature-complete API for some programming languages (like Python and JavaScript), it doesn't do so for others like C++ which is only supported for clients. While there are libraries to create C++ WebSocket servers and clients (like websocket++ which is used in this paper), they require a lot of code (to some part due to the nature of the language) and tedious state management when used directly. Additionally, the event system would need to be implemented manually for every event every time, which is tedious. \cite{b4}
    \item The data serialization and deserialization of Socket.IO is fully dynamic, meaning that before an event is actually processed, the code cannot possibly know the structure of the event data. This presents multiple problems:
    \begin{itemize}
        \item Deserialization of event data to language-native data structures is impossible for statically typed languages like C++, which negates the benefits of static typing. It requires the use of dynamic containers like maps to represent objects with strings as attribute names. Since the containers need to be fully dynamic, the programmer first needs to write code to check if an expected attribute exists, assert its's type and finally convert it to the representative language-native type, which is tedious.
        \item Even in dynamic languages like JavaScript where objects can have dynamic attributes and variables are dynamically typed it is still helpful to perform the same validation in order to provide adequate error handling and avoid unexpected behaviors. There are also additional systems (TypeScript for JS and type hints for Python) that provide the benefits of static typing (such as rich IntelliSense) even though the underlying language is dynamic.
    \end{itemize}
\end{enumerate}

@@

Similar to the protocols and accompanying libraries listed above (especially Socket.IO), msglink tries to achieve it's core functionality with the least code possible using a declarative API. Rather than writing code explicitly stating how to perform the serialization, validation or communication step-by-step (imperative programming), the developer writes code specifying (declaring) what result is expected (declarative programming). \cite{imperative-vs-declarative}

@@

Since msglink uses WebSockets as an underlying transport layer, a rather inefficient JSON encoding for serialization and is an internally rather complex protocol to manage, speed is not one of it's strengths. As mentioned in the protocol stackup, directly using a more lower-level protocol like TCP or UDP is required for high-performance applications, trading convenience for speed.

@@

msglink satisfies all the goals defined in this paper for it's specific use case and allows building new, innovative applications quicker without worrying about communication.

>>>>

\begin{enumerate}
    \item While the event system provided by Socket.IO is useful for many applications, sometimes responses to events are required. This is typically named a \textbf{Remote Procedure Call} and needs to be implemented manually for every event:
    \begin{itemize}
        \item define and keep track of a separate request and response event
        \item add an identifying element to request and response data to identify which response belongs to which request and prevent mixing up responses if multiple requests are sent out in short succession
        \item keep a list of open requests to handle responses once received.
    \end{itemize}
    \item While Socket.IO provides a very clean and feature-complete API for some programming languages (like Python and JavaScript), it doesn't do so for others like C++ which is only supported for clients. While there are libraries to create C++ WebSocket servers and clients (like websocket++ which is used in this paper), they require a lot of code (to some part due to the nature of the language) and tedious state management when used directly. Additionally, the event system would need to be implemented manually for every event every time, which is tedious. \cite{b4}
    \item The data serialization and deserialization of Socket.IO is fully dynamic, meaning that before an event is actually processed, the code cannot possibly know the structure of the event data. This presents multiple problems:
    \begin{itemize}
        \item Deserialization of event data to language-native data structures is impossible for statically typed languages like C++, which negates the benefits of static typing. It requires the use of dynamic containers like maps to represent objects with strings as attribute names. Since the containers need to be fully dynamic, the programmer first needs to write code to check if an expected attribute exists, assert its type and finally convert it to the representative language-native type, which is tedious.
        \item Even in dynamic languages like JavaScript where objects can have dynamic attributes and variables are dynamically typed it is still helpful to perform the same validation in order to provide adequate error handling and avoid unexpected behaviors. There are also additional systems (TypeScript for JS and type hints for Python) that provide the benefits of static typing (such as rich IntelliSense) even though the underlying language is dynamic.
    \end{itemize}
\end{enumerate}

@@

Similar to the protocols and accompanying libraries listed above (especially Socket.IO), msglink tries to achieve its core functionality with the least code possible using a declarative API. Rather than writing code explicitly stating how to perform the serialization, validation or communication step-by-step (imperative programming), the developer writes code specifying (declaring) what result is expected (declarative programming). \cite{imperative-vs-declarative}

@@

Since msglink uses WebSockets as an underlying transport layer, a rather inefficient JSON encoding for serialization and an internally rather complex protocol to manage, speed is not one of its strengths. As mentioned in the protocol stackup, directly using a more lower-level protocol like TCP or UDP is required for high-performance applications, trading convenience for speed.

@@

msglink satisfies all the goals defined in this paper for its specific use case and allows building new, innovative applications quicker without worrying about communication.
